---
title: "Test 2"
output:
  html_notebook: default
  pdf_document: default
fontsize: 9pt
---


Loading the packages first
```{r setoptions, echo=TRUE }

library(pacman)
    p_load(tidyverse, lavaan, psych, knitr, lme4, lmerTest, multilevel, nlme, lattice, sjPlot, ggplot2,cowplot, magrittr, broom, Metrics, pbkrtest, mlmRev, influence.ME, gridExtra, semPlot)
    
```
Question 1 – Structural Equation Modeling

loading data
```{r loaddata, mysize = TRUE, message = FALSE, size="\\scriptsize"}

data <- readRDS("peru2.RDS")  

```

Making variables for the path analysis for SEM. I would like to predict risky/antisocial behaviour using self-efficacy, early life SES (round 1 and 2), mental health and agency
```{r}

#   SES (Using the following variables: Housing quality, Access to services, Consumer durables)
    data$hq_r12<-data$hq_r12/3
    data$sv_r12<-data$sv_r12/3
    data$cd_r12<-data$cd_r12/3
    data$ses_scale<-data$hq_r12+data$sv_r12+data$cd_r12
    
#   self-efficacy
    data$self_eff1<-(data$self_eff1)/8    
    data$self_eff2<-(data$self_eff2)/8  
    data$self_eff3<-(data$self_eff3)/8
    data$self_eff4<-(data$self_eff4)/8
    data$self_eff5<-(data$self_eff5)/8
    data$self_eff6<-(data$self_eff6)/8
    data$self_eff7<-(data$self_eff7)/8
    data$self_eff8<-(data$self_eff8)/8
    data$self_efficacy_scale<-(data$self_eff1 + data$self_eff2 + data$self_eff3 + data$self_eff4 + data$self_eff5 + data$self_eff6 + data$self_eff7 + data$self_eff8)
    
#   agency
    data$agency1<-data$agency1/4
    data$agency2<-data$agency2/4
    data$agency3<-data$agency3/4
    data$agency4<-data$agency4/4
    data$agency_scale<-data$agency1+data$agency2+data$agency3+data$agency4
    
#   mental wellness
    data$sdq1<-data$sdq1/5
    data$sdq2<-data$sdq2/5
    data$sdq3<-data$sdq3/5
    data$sdq4<-data$sdq4/5
    data$sdq5<-data$sdq5/5
    data$mental_wellness_scale<-data$sdq1+data$sdq2+data$sdq3+data$sdq4+data$sdq5
    
#   anti-social, or risk-taking behaviour at Round 5 
    
    data$FRNSMKR5<-data$FRNSMKR5/9 #Have friends who smoke   
    data$FRNALCR5<-data$FRNALCR5/9 #Have friends who use alcohol
    data$YOUALCR5<-data$YOUALCR5/9 #Uses alcohol 
    data$BEATEN<-data$BEATEN/9     #Beaten up by friends, strangers, teachers, parents 
    data$ARRSTDR5<-data$ARRSTDR5/9 #Has been arrested
    data$FRNGNGR5<-data$FRNGNGR5/9 #Has friends in gang
    data$MEMGNGR5<-data$MEMGNGR5/9 #Is a member of a gang
    data$CRYWPNR5<-data$CRYWPNR5/9 #Has carried a weapon
    data$NUMPRTR5<-data$NUMPRTR5/9 #Number of sex partners had
    data$risky_behaviour_scale<-(data$FRNSMKR5 + data$FRNALCR5 + data$YOUALCR5 + data$BEATEN + data$ARRSTDR5 + data$FRNGNGR5 + data$MEMGNGR5 + data$CRYWPNR5 + data$NUMPRTR5)
```
 Defining path model
```{r}

pathm1 <- '
      risky_behaviour_scale ~ self_efficacy_scale
      mental_wellness_scale ~ self_efficacy_scale

'
prejpathfit1 <- sem(pathm1, data = data)
summary(prejpathfit1, fit.measures = T) 

```
 
CFI and TLI are > .09 which shows that the model has a good fit, The model is  shows a good fit, not too sure about the RMSEA, p-value =Na, might need to compare to another model


Now drawing model using package semPlot
```{r}
semPlot::semPaths(prejpathfit1, what = "est", layout = "spring")
```
Our model is telling us that self efficacy and mental wellness are predictors of risky behaviours with self-efficacy also predicting mental illness

Defining path model 2

```{r predicttree, mysize = TRUE, warning=FALSE,  size="\\scriptsize", fig.height=3}

pathm2 <- '
      risky_behaviour_scale ~ self_efficacy_scale + mental_wellness_scale
      mental_wellness_scale ~ self_efficacy_scale + agency_scale
      self_efficacy_scale ~  agency_scale 
'    
prejpathfit2 <- sem(pathm2, data = data)
summary(prejpathfit2, fit.measures = T)   
anova(prejpathfit2, prejpathfit1) #comparing this model to our first model
```

Now plotting this model
```{r}
semPlot::semPaths(prejpathfit2, what = "est", layout = "spring")
```
Let's take a closer look at the residuals as our new model as the AIC is higher than the first model's one and the RMSEA p-value is not significant
```{r}
      resid(prejpathfit2, type = "normalized")
      modificationindices(prejpathfit2)

```
we can see correlation between the risky behaviour scale and the mental wellness and self-efficacy scales
now lets respecify the model to allow correlations between the error terms in each scale 
```{r}
  
  prejpathmodel2 <- '
                # MEASUREMENT MODEL
                  agency_scale =~ agency1 + agency2 + agency3 + agency4
                  self_efficacy_scale =~ self_eff1 + self_eff2 + self_eff3 + self_eff4 + self_eff5 + self_eff6                   + self_eff7 + self_eff8
                  mental_wellness_scale =~ sdq1 + sdq2 + sdq3 + sdq4 + sdq5
                  risky_behaviour_scale =~ FRNSMKR5 + FRNALCR5 + YOUALCR5 + BEATEN + ARRSTDR5 + FRNGNGR5 +                      MEMGNGR5 + CRYWPNR5 + NUMPRTR5

      
                # STRUCTURAL MODEL
                  risky_behaviour_scale ~ self_efficacy_scale + mental_wellness_scale
                  mental_wellness_scale ~ self_efficacy_scale + agency_scale
                  self_efficacy_scale ~  agency_scale 
                  
                # CORRELATED ERRORS
                  risky_behaviour_scale	~~	mental_wellness_scale
                  risky_behaviour_scale	~~	self_efficacy_scale
      
               '
 prejpathfit2 <- sem(prejpathmodel2, data = data)
      summary(prejpathfit2, fit.measures = T)

```



